### 2020-09-16 paper/appendA.tex
many many fixes

\section{Introduction}
\TODO{some intro}
We give a brief overview of the square-free perfect graph coloring algorithm and a pseudocode thereof. The work is not intended to be self-contained and will be impossible to fully understand without \cite{coloringSquareFree} at hand. But when reading \cite{coloringSquareFree}, one may pose a question of what sort of an algorithm emerges from it, and we believe the pseudocode below gives a first overview and is a major step required before an attempt to implement it would be made. For some, looking at the pseudocode could provide a better way to analyze and possibly simplify the algorithm.

@@

In 2003 Maffray et al. published a paper on Artemis graphs where they show polynomial algorithms to recognize and color them \cite{Maffray2006}. In 2005 Maffray et al. published a paper on Artemis graphs\cite{Maffray2005} they showed an polynomial algorithm that checks if a perfect graph contains a prism, and returns it if so.

@@

\begin{defnTwo}[$J$-strip system, $uv$-rung]
  Let $J$ be a 3-connected graph and let $G$ be a perfect graph. A \emph{$J$-strip system} $(S, N)$ in $G$ means
  \begin{itemize}
    \item for each edge $uv$ of $J$, a subset $S_{uv} = S_{vu}$ of $V(G)$,
    \item for each vertex $v$ of $J$, s subset $N_v$ of $V(G)$,
    \item $N_{uv} = S_{uv} \cap N_u$,
  \end{itemize}
  such that if we define a $uv$-rung to be a path $R$ of $G$ with ends $s, t$, where $V(R) \subseteq S_{uv}$, and $s$ is the unique vertex of $R$ in $N_u$, and $t$ is the unique vertex of $R$ in $N_v$, the following conditions are met:
  \begin{itemize}
    \item the sets $S_{uv}$, for $uv \in E(J)$ are pairwise disjoint,
    \item for each $u \in V(J)$, $N_u \subseteq \bigcup_{uv \in E(J)} S_{uv}$,
    \item for each $uv \in E(J)$, every vertex of $S_{uv}$ is in a $uv$-rung,
    \item for any two edges $uv, wx$ of $J$, with $u, v, w, x$ all distinct, there are no edges between $S_{uv}$ and $S_{wx}$,
    \item if $uv, uw \in E(J)$, with $v \neq w$, then $N_{uv}$ is complete to $N_{uw}$ and there are no other edges between $S_{uv}$ and $S_{uw}$,
    \item for each $uv \in E(J)$, there is a \emph{special} $uv$-rung, such that for every cycle $C$ of $J$, the sum of the lengths of the special $uv$-rungs for $uv \in E(C)$ has the same parity as $|V(C)|$.
  \end{itemize}
\end{defnTwo}

@@

A $J$-strip system has the following properties:
\begin{itemize}
  \item for distinct $u, v \in V(J)$, if $uv \in E(J)$, then $N_u \cap N_v \subseteq S_{uv}$ and if $uv \notin E(J)$ then $N_u \cap N_v = \emptyset$,
  \item for $uv \in E(J)$ and $w \in V(J)$, if $w \neq u$ then $S_{uv} \cap N_w = \emptyset$,
  \item for every $uv \in E(J)$, all $uv$-rungs have lengths of the same parity,
  \item for every cycle $C$ of $J$ and every choice of $uv$-rung for every $uv \in E(C)$, the sums of the lengths of the $uv$-rungs have the same parity as $|V(C)|$. In particular, for each edge $uv \in E(J)$, all $uv$-rungs have the same parity.
\end{itemize}

@@

\begin{itemize}
	\item $a :\in X$, when $X$ is a set -- let a be equal to any element of $X$,
	\item $a \xor b$, when $a$ and $b$ are logical values -- $a$ xor $b$,
	\item $V(X)$ -- vertices of structure $X$. Will be written as $X$ when obvious,
	\item $a - b$, when $a$ and $b$ are nodes -- $a$ and $b$ are neighbors,
	\item $a \cdots b$, when $a$ and $b$ are nodes -- $a$ and $b$ are not neighbors,
	\item $a - X$, when $a$ is a node and $X$ is a set of nodes -- $a$ has a neighbor in $X$,
	\item $a \cdots X$, when $a$ is a node and $X$ is a set of nodes -- $a$ has a nonneighbor in $X$,
	\item $a \blacktriangleleft  X$, when $a$ is a node and $X$ is a set of nodes -- $a$ is complete to $X$,
	\item $a \ntriangleleft X$, when $a$ is a node and $X$ is a set of nodes -- $a$ is anticomplete to $X$,
	\item $X \setComplete Y$, when $X$ and $Y$ are set of nodes -- $X$ is complete to $Y$,
	\item $X \setAntiComplete Y$, when $X$ and $Y$ are set of nodes -- $X$ is anticomplete to $Y$,
	\item $[n]$  -- $\{1, \ldots, n\}$,
	\item \LGBSK -- the line-graph of a biparite subdivision of $K_4$.
\end{itemize}

>>>>

\section{Introduction}
We give a brief overview of the square-free perfect graph coloring algorithm and a pseudocode thereof. The work is not intended to be self-contained and will be impossible to fully understand without \cite{coloringSquareFree} at hand. But when reading \cite{coloringSquareFree}, one may pose a question of what sort of an algorithm emerges from it, and we believe the pseudocode below gives a first overview and is a major step required before an attempt to implement it would be made. For some, looking at the pseudocode could provide a better way to analyze and possibly simplify the algorithm.

@@

In 2003 Maffray et al. published a paper on Artemis graphs where they show polynomial algorithms to recognize and color them \cite{Maffray2006}. In 2005 Maffray et al. published a paper on Artemis graphs \cite{Maffray2005} they showed an polynomial algorithm that checks if a perfect graph contains a prism, and returns it if so.

@@

\begin{defnTwo}[$J$-strip system, $uv$-rung]
  Let $J$ be a 3-connected graph and let $G$ be a perfect graph. A \emph{$J$-strip system} $(S, N)$ in $G$ means
  \begin{itemize}
    \item for each edge $uv$ of $J$, a subset $S_{uv} = S_{vu}$ of $V(G)$,
    \item for each vertex $v$ of $J$, s subset $N_v$ of $V(G)$,
    \item $N_{uv} = S_{uv} \cap N_u$,
  \end{itemize}
  such that if we define a $uv$-rung to be a path $R$ of $G$ with ends $s, t$, where $V(R) \subseteq S_{uv}$, and $s$ is the unique vertex of $R$ in $N_u$, and $t$ is the unique vertex of $R$ in $N_v$, the following conditions are met:
  \begin{itemize}
    \item the sets $S_{uv}$, for $uv \in E(J)$ are pairwise disjoint,
    \item for each $u \in V(J)$, $N_u \subseteq \bigcup_{uv \in E(J)} S_{uv}$,
    \item for each $uv \in E(J)$, every vertex of $S_{uv}$ is in a $uv$-rung,
    \item for any two edges $uv, wx$ of $J$, with $u, v, w, x$ all distinct, there are no edges between $S_{uv}$ and $S_{wx}$,
    \item if $uv, uw \in E(J)$ with $v \neq w$, then $N_{uv}$ is complete to $N_{uw}$ and there are no other edges between $S_{uv}$ and $S_{uw}$,
    \item for each $uv \in E(J)$, there is a \emph{special} $uv$-rung, such that for every cycle $C$ of $J$, the sum of the lengths of the special $uv$-rungs for $uv \in E(C)$ has the same parity as $|V(C)|$.
  \end{itemize}
\end{defnTwo}

@@

A $J$-strip system has the following properties:
\begin{itemize}
  \item for distinct $u, v \in V(J)$, if $uv \in E(J)$, then $N_u \cap N_v \subseteq S_{uv}$ and if $uv \notin E(J)$, then $N_u \cap N_v = \emptyset$,
  \item for $uv \in E(J)$ and $w \in V(J)$, if $w \neq u$, then $S_{uv} \cap N_w = \emptyset$,
  \item for every $uv \in E(J)$, all $uv$-rungs have lengths of the same parity,
  \item for every cycle $C$ of $J$ and every choice of $uv$-rung for every $uv \in E(C)$, the sums of the lengths of the $uv$-rungs have the same parity as $|V(C)|$. In particular, for each edge $uv \in E(J)$, all $uv$-rungs have the same parity.
\end{itemize}

@@

\begin{itemize}
	\item $a :\in X$, when $X$ is a set -- let $a$ be equal to any element of $X$,
	\item $a \xor b$, when $a$ and $b$ are logical values -- $a$ xor $b$,
	\item $V(X)$ -- vertices of structure $X$. Will be written as $X$ when obvious,
	\item $a - b$, when $a$ and $b$ are nodes -- $a$ and $b$ are neighbors,
	\item $a \cdots b$, when $a$ and $b$ are nodes -- $a$ and $b$ are not neighbors,
	\item $a - X$, when $a$ is a node and $X$ is a set of nodes -- $a$ has a neighbor in $X$,
	\item $a \cdots X$, when $a$ is a node and $X$ is a set of nodes -- $a$ has a nonneighbor in $X$,
	\item $a \blacktriangleleft  X$, when $a$ is a node and $X$ is a set of nodes -- $a$ is complete to $X$,
	\item $a \ntriangleleft X$, when $a$ is a node and $X$ is a set of nodes -- $a$ is anticomplete to $X$,
	\item $X \setComplete Y$, when $X$ and $Y$ are set of nodes -- $X$ is complete to $Y$,
	\item $X \setAntiComplete Y$, when $X$ and $Y$ are set of nodes -- $X$ is anticomplete to $Y$,
	\item $[n]$  -- $\{1, \ldots, n\}$,
	\item \LGBSK -- the line-graph of a biparite subdivision of $K_4$.
\end{itemize}

### 2020-09-16 paper/coloringBergeGraphs.tex
Poprawki up to old Ch 1
many many fixes

The polynomial technique of coloring perfect graphs known so far arose in the field of semidefinite programming. Semidefinite programs are linear programs over the cones of semi-definite matrices. The connection of coloring graphs and the cones of semi-definite matrices might be surprising, so let us take a brief digression into the field of information theory, where we will see the connection more clearly. Also, this was exactly the background which motivated Berge to introduce perfect graphs \cite{Chudnovsky2003}.

\subsection{Shannon Capacity of a graph}
\label{sec:ShannonCapacity}

@@

Consider an "umbrella" in $\mathbb{R}^3$ with the unit vector $e_1 = (1, 0, 0)$ as its "handle" and 5 "ribs" of unit length.  Open it up to the point where non-consecutive ribs are orthogonal, that is form an angle of 90$^\circ$. This way we get a representation of $C_5$ by 5 unit vectors $u_1, \ldots u_5$ so that each $u_i$ forms the same angle with $e_1$ and any two non-adjacent nodes are represented with orthogonal vectors. See \Cref{fig:umbrella}. We can calculate $e_1^\intercal u_i = 5 ^ {-1/4}$.

\begin{wrapfigure}{l}{0.4\textwidth}
  \input{tikzpictures/umbrella.tex}
  \caption{Lovász's umbrella}
  \label{fig:umbrella}
\end{wrapfigure}

@@

If $S$ is any stable set in $C_5^k$, then $\{v_i, i \in S\}$ is a set of mutually orthogonal unit vectors so we get \todo{why?}
$$\sum\limits_{i\in S}(e_1^\intercal v_i)^2 \leq |e_1|^2 = 1$$
(if $v_i$ formed a basis then this inequality would be an equality).

On the other hand each term on the left hand side is $5^{-1/4}$, so the left hand side is equal to $|S|5^{-k/2}$, and so $|S| \leq 5^{k/2}$. Since $|S|$ was an arbitrary stable set, we get $\alpha(C_5^k) \leq 5 ^{k/2}$ and $\Theta(C_5) = \sqrt{5}$.

It turns out that this method extends to any graph $G$ in place of $C_5$. All we have to do is find a orthogonal representation that will give us the best bound. So, we can define the \emph{Lovász number} of a graph $G$ as \todo{this equation does not really follow from the thought process above, it is a slightly different definition}:
$$\vartheta(G) = \min\limits_{c,U} \max\limits_{i\in V} \frac{1}{(c^\intercal u_i)^2},$$

\noindent where $c$ is a unit vector in $\mathbb{R}^{|V(G)|}$ and $U$ is a orthogonal representation of $G$.

Contrary to Lovász's first hope \cite{Lovasz1979} $\vartheta(G)$ does not always equal $\Theta(G)$, it is only an upper bound on it. However, these two are equal for some graphs, including all perfect graphs, as is demonstrated in the Lovász "sandwich theorem".

\begin{theorem}[Lovász "sandwich theorem" \cite{Knuth1994}]
  \label{thm:sandwich}
  For any graph $G$:
  $$ \omega(G) \leq \vartheta(\overline{G}) \leq \chi(G).$$
\end{theorem}

@@

But how can we construct an optimum (or even good) orthogonal representation? It turns out that it can be computed in polynomial time using semidefinite optimization.

@@

\begin{defn}[eigenvector, eigenvalue]
  Let $A$ be an $n \times n$ real matrix. An \emph{eigenvector} of $A$ is a vector $x$ such that $Ax$ is parallel to $x$. In other words, there is a real or complex number $\lambda$, such that $Ax = \lambda x$. This $\lambda$ is called the \emph{eigenvalue} of $A$ belonging to eigenvector $x$.
\end{defn}

@@

Solving semidefinite programs is a complex topic, we refer to \cite{grotschel1993} for reference. All we need to know is that we can solve semidefinite programs up to an arbitrarily small error in polynomial time. One of the methods to do this is called the \emph{ellipsoid method}, hence the name for the coloring algorithm.

@@

\section{Coloring perfect graph using ellipsoid method}
\label{sec:coloringEllipsoid}
The following is based on \citetitle{Laurent2005} by \citeauthor{Laurent2005} \cite{Laurent2005}.

@@

Given a natural weight function $w : V \rightarrow \mathbb{N}$ we could calculate the maximum weighted stable set in $G$ using the following modification:

@@

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$. We will call recursion at most $O(|V|)$ times, each step of recursion is polynomial in time. Therefore the running time of \Cref{alg:minColoring} is polynomial.

@@

\begin{defn}[prism]
  A \emph{prism} is a graph consisting of two disjoint triangles and two disjoint paths between them. Notice, that for a graph to contain no odd hole, all three paths in a prism must have the same parity. A prism with all three paths odd is called an \emph{odd prism}.
  \label{def:prism}
\end{defn}

In 2005 Maffray and Trotignon a coloring algorithm that colors graphs containing no odd hole, no antihole and no prism (sometimes called Artemis graphs) in $O(|V|^4|E|)$ time \cite{Maffray2006}. They later improved the time complexity to $O(|V|^2|E|)$ \cite{Lvque2009}.

@@

In 2016 Chudnovsky et al. published an algorithm that given a perfect graph $G$ with $\omega(G) = k$ colors it optimally in a time polynomial for a fixed $k$ \cite{Chudnovsky2017}.

A most recent advancement (2018) is an algorithm by Chudnovsky et al. that colors any square-free Berge graphs in time of $O(|V|^9)$ \cite{Chudnovsky2019}. Before proving strong perfect graph conjecture, a similar conjecture for square-free Berge graphs has been proven by Conforti et al. \cite{Conforti2004}. During one of her lectures, Maria Chudnovsky expressed hope that discovery of full algorithm for coloring Berge graphs might follow a similar pattern.

The algorithm for coloring square-free Berge graphs is much more complex than CCLSV algorithm for Berge graphs recognition and would be much harder to implement. The main idea is that, if square-free Berge graph $G$ is a member of class that is simpler to analyze and color (Artemis graphs), then it must have an induced prism. We find it and grow it until it is maximal and then we can create a partition of $G$, that we color recursively. The paper in which the algorithm is presented \cite{Chudnovsky2019} is of much more mathematical ''flavor'' then \cite{MC05} -- there is a proof that a $O(|V|^9)$ algorithm exists, but actually deducing it from the paper is far from straightforward. We do it in the Appendix A.

>>>>

The polynomial technique of coloring perfect graphs known so far arose in the field of semidefinite programming. Semidefinite programs are linear programs over the cones of semidefinite matrices. The connection of coloring graphs and the cones of semidefinite matrices might be surprising, so let us take a brief digression into the field of information theory, where we will see the connection more clearly. Also, this was exactly the background which motivated Berge to introduce perfect graphs \cite{Chudnovsky2003}.

\subsection{Shannon capacity of a graph}
\label{sec:ShannonCapacity}

@@

Consider an "umbrella" in $\mathbb{R}^3$ with the unit vector $e_1 = (1, 0, 0)$ as its "handle" and 5 "ribs" of unit length.  Open it up to the point where non-consecutive ribs are orthogonal, i.e., form an angle of 90$^\circ$. This way we get a representation of $C_5$ by 5 unit vectors $u_1, \ldots u_5$ so that each $u_i$ forms the same angle with $e_1$ and any two non-adjacent nodes are represented with orthogonal vectors. See \Cref{fig:umbrella}. We can calculate $e_1^\intercal u_i = 5 ^ {-1/4}$.

\begin{wrapfigure}{r}{0.4\textwidth}
  \input{tikzpictures/umbrella.tex}
  \caption{Lovász's umbrella}
  \label{fig:umbrella}
\end{wrapfigure}

@@

If $S$ is any stable set in $C_5^k$, then $\{v_i: i \in S\}$ is a set of mutually orthogonal unit vectors, so we could extend $S$ to $S'$, so that $\{v_i: i \in S'\}$ is a basis of $\mathbb{R}^{3k}$. This gives us
$$\sum\limits_{i\in S}(e_1^\intercal v_i)^2 \leq \sum\limits_{i \in S'}(e_1^\intercal v_i)^2 \leq ? = 1$$
\todo{todo}
Each term on the left hand side is $5^{-k/4}$, so the left hand side is equal to $|S|5^{-k/2}$, and so $|S| \leq 5^{k/2}$. Since $|S|$ was an arbitrary stable set, we get $\alpha(C_5^k) \leq 5 ^{k/2}$ and $\Theta(C_5) = \sqrt{5}$.

It turns out that this method extends to any graph $G$ in place of $C_5$. All we have to do is to find a orthogonal representation that will give us the best bound. So, we can define the \emph{Lovász number} of a graph $G$ as \todo{this equation does not really follow from the thought process above, it is a slightly different definition}:
$$\vartheta(G) = \min\limits_{c,U} \max\limits_{i\in V} \frac{1}{(c^\intercal u_i)^2},$$

\noindent where $c$ is a unit vector in $\mathbb{R}^{|V(G)|}$ and $U$ is an orthogonal representation of $G$.

Contrary to Lovász's first hope \cite{Lovasz1979} $\vartheta(G)$ does not always equal $\Theta(G)$, but it is only an upper bound on it. However, these two are equal for some graphs, including all perfect graphs, as is demonstrated in the Lovász "Sandwich theorem".

\begin{theorem}[Lovász "Sandwich theorem" \cite{Knuth1994}]
  \label{thm:sandwich}
  For any graph $G$:
  $$ \omega(G) \leq \vartheta(\overline{G}) \leq \chi(G).$$
\end{theorem}

@@

But how can we construct an optimum (or even sufficiently good) orthogonal representation? It turns out that it can be computed in polynomial time using semidefinite optimization.

@@

\begin{defn}[eigenvector, eigenvalue]
  Let $A$ be an $n \times n$ real matrix. An \emph{eigenvector} of $A$ is a vector $x$ such that $Ax$ is parallel to $x$. In other words, there is a real or a complex number $\lambda$, such that $Ax = \lambda x$. This $\lambda$ is called the \emph{eigenvalue} of $A$ belonging to eigenvector $x$.
\end{defn}

@@

Solving semidefinite programs is a complex topic, we refer to \cite{grotschel1993} for details. All we need to know is that we can solve semidefinite programs up to an arbitrarily small error in polynomial time. One of the methods to do this is called the \emph{ellipsoid method}, hence the name for the coloring algorithm.

@@

\section{Coloring perfect graph using ellipsoid method}
\label{sec:coloringEllipsoid}
The following part is based on \citetitle{Laurent2005} by \citeauthor{Laurent2005} \cite{Laurent2005}.

@@

Given a weight function $w : V \rightarrow \mathbb{N}$ we could calculate the maximum weighted stable set in $G$ using the following modification:

@@

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$. We will call recursion at most $O(|V|)$ times, each step of recursion requires polynomial time. Therefore the total running time of \Cref{alg:minColoring} is polynomial.

@@

\begin{defn}[prism]
  A \emph{prism} is a graph consisting of two disjoint triangles and two disjoint paths between them. Notice that for a graph to contain no odd hole, all three paths in a prism must have the same parity. A prism with all three paths odd is called an \emph{odd prism}.
  \label{def:prism}
\end{defn}

In 2005 Maffray and Trotignon a coloring algorithm that colors graphs containing no odd hole, no antihole and no prism (sometimes called Artemis graphs) in $O(|V|^4|E|)$ time \cite{Maffray2006}. They later improved the time complexity to $O(|V|^2|E|)$ \cite{coloringArtemis}.

@@

In 2016 Chudnovsky et al. published an algorithm that given a perfect graph $G$ with $\omega(G) = k$ colors it optimally in a time polynomial for a fixed $k$ \cite{Chudnovsky2017}. In general case this algorithm has a time complexity of $O(|V|^{(k+1)^2})$.

A most recent advancement (2018) is an algorithm by Chudnovsky et al. that colors any square-free Berge graphs in time of $O(|V|^9)$ \cite{coloringSquareFree}. Before proving strong perfect graph conjecture, a similar conjecture for square-free Berge graphs has been proven by Conforti et al. \cite{Conforti2004}. During one of her lectures, Maria Chudnovsky expressed hope that discovery of full algorithm for coloring Berge graphs might follow a similar pattern.

The algorithm for coloring square-free Berge graphs is much more complex than CCLSV algorithm for Berge graphs recognition and would be much harder to implement. The main idea is that, if square-free Berge graph $G$ is a member of class that is simpler to analyze and color (Artemis graphs), then it must have an induced prism. We find it and grow it until it is maximal and then we can create a partition of $G$, that we color recursively. The paper in which the algorithm is presented \cite{coloringSquareFree} is of much more mathematical ''flavor'' than \cite{MC05} -- there is a proof that a $O(|V|^9)$ algorithm exists, but actually deducing it from the paper is far from straightforward. We provide its pseudocode in the Appendix A.

### 2020-09-16 paper/experiments.tex
many many fixes

We ran tests on nine classes of perfect graphs.
\begin{itemize}
  \item perfect Erd\H{o}s--Rényi graphs,
  \item Random bipartite graphs,
  \item Line graphs of random bipartite graphs,
  \item Lattice graphs (\cref{fig:lattice}),
  \item Rook graphs (\cref{fig:rook}),
  \item Knight graphs (\cref{fig:knight}),
  \item Hypercube graphs (\cref{fig:hypercube}),
  \item Split graphs (\cref{fig:split}),
  \item Full binary trees.
\end{itemize}

@@

We generate random bipartite graphs by creating two equal-sized sets of vertices and every possible edge between them has a probability of $1/2$ to appear. We also generate line graphs of bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$.

@@

Let us turn our attention to the GPU CCLSV's results. \Cref{plot:perfDetailed} shows breakdown of each implementation running times on a single graph of each class. We can clearly see that in most cases, testing possible near-cleaners takes most of the running time, so the speedup by the GPU is significant.

Examples of full binary graph, and lattice graph show that getting near cleaners still can be a big bottleneck. Parallelization of this method could be considered in the future. A possible idea would be to use recently developed dynamic dictionary for the GPU \cite{Ashkiani2018}. When running GPU CCLSV on split graphs, quite a lot of time is spent checking simple structures. This should be easily parallelizable if need be, or if graphs that require it are encountered.

@@

\section{Coloring Berge Graphs}

@@

First, we present overall running times of the coloring algorithm (\Cref{fig:CSDPLines}) and then breakdown the running time in two parts: calculating $\vartheta$ by the CSDP library and graphs manipulations which constitute the rest of the coloring algorithm (\Cref{fig:CSDPDet}). We note that $\vartheta$ calculation is highly dependent on the type of the graph. For our biggest tested cases of $|V| = 48$, all required calculations of $\vartheta$ take almost 7 minutes in biparite graphs, but only 12 seconds in line graphs of bipartite graphs. Because CSDP library is highly optimized, it is expected that it would run much faster on some graphs. Hovewer, it is not clear what could we do to speedup the running times on tests that CSDP is slow on, such as bipartite graphs of lattice graphs. A potential improvement would be to again use CUDA, this time for $\vartheta$ calculations, possibly utilizing the cuSOLVER library \cite{cusolver}.

We notice that on some graphs, namely line graphs of bipartite graphs, rook graphs and split graphs the time taken by operations other than calculating $\vartheta$ by the CSDP library is quite significant. Hovewer, the overall running times on these graphs are low, so we didn't particularly optimize this, so probably some improvements could be made if need be.

@@

>>>>

We ran tests on eight classes of perfect graphs:
\begin{itemize}
  % \item perfect Erd\H{o}s--Rényi graphs,
  \item random bipartite graphs,
  \item line graphs of random bipartite graphs,
  \item lattice graphs (\cref{fig:lattice}),
  \item rook graphs (\cref{fig:rook}),
  \item knight graphs (\cref{fig:knight}),
  \item hypercube graphs (\cref{fig:hypercube}),
  \item split graphs (\cref{fig:split}),
  \item full binary trees.
\end{itemize}

\todo{generate perfect Erdos--Renyi graphs}

@@

We generate random bipartite graphs by creating two equal-sized sets of vertices and every possible edge between them has a probability of $p = 0.5$ to appear. In this way we can also generate line graphs of bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$.

@@

Let us turn our attention to the GPU CCLSV's results. \Cref{plot:perfDetailed} shows breakdown of each implementation running times on a single graph of each class. We can clearly see that in most cases testing possible near-cleaners takes most of the running time, so the speedup by the GPU is significant.

Examples of full binary graphs, and lattice graphs show that getting near cleaners still can be a big bottleneck. Parallelization of this method could be considered in the future. A possible idea would be to use recently developed dynamic dictionary for the GPU \cite{Ashkiani2018}. When running GPU CCLSV on split graphs, quite a lot of time is spent checking simple structures. This should be parallelizable, the biggest challange we can see is spreading work evenly across CUDA threads.

@@

\section{Coloring Berge graphs}

@@

@@

First, we present overall running times of the coloring algorithm (\Cref{fig:CSDPLines}) and then breakdown the running time in two parts: calculating $\vartheta$ by the CSDP library and graphs manipulations which constitute the rest of the coloring algorithm (\Cref{fig:CSDPDet}). We note that $\vartheta$ calculation is highly dependent on the type of the graph. For our biggest tested cases of $|V| = 48$, all required calculations of $\vartheta$ take almost 7 minutes in biparite graphs, but only 12 seconds in line graphs of bipartite graphs. Because CSDP library is highly optimized, it is expected that it would run much faster on some graphs. However, it is not clear what could we do to speedup the running times on tests that CSDP is slow on, such as bipartite graphs of lattice graphs. A potential improvement would be to again use CUDA, this time for $\vartheta$ calculations, possibly utilizing the cuSOLVER library \cite{cusolver}.

We notice that on some graphs, namely line graphs of bipartite graphs, rook graphs and split graphs the time taken by operations other than calculating $\vartheta$ by the CSDP library is quite significant. However, the overall running times on these graphs are low, so we didn't particularly optimize this, so probably some improvements could be made here.

\section{Conclusions}
We implemented and benchmarked two different approaches for recognizing perfect graphs. Both have their positives and drawbacks -- CCLSV's running times are more predictable, but on many graphs it is noticeably slower than the na\"ive algorithm. On the other hand, running na\"ive algorithm on a graph we don't know any properties of can lead to unbearably slow running time.

The GPU CCLSV implementation shows that using parallel computation can be very profitable for this task. A further research can be done to see if further progress could be made, however there is no obvious improvement we can see. Generating all possible near-cleaners makes a heavy use of sets (our tests indicated that if we simply list all combinations of sets $X_i$ and $N_j$, the resulting list $\mathcal{R}$ would be much bigger than necessary). Possibly a newly developed dynamic dictionary for the GPU \cite{Ashkiani2018} would be of use here. Testing for simple forbidden structures should also be parallelizable, although our preliminary tests indicated that spreading all the work evenly between CUDA threads is not trivial.

A GPU implementation of na\"ive algorithm could also be considered for future work. We have created one, but after all path-enumaration optimizations of the CPU implementation we didn't manage to achieve any noticeable speedup. The reason again was, that we could't spread the work evenly between CUDA threads. Because some prefixes of paths require much more processing than others, this is not a trivial task and a careful study would be needed.

We also implemented algorithm for coloring perfect graphs. It is not entirely unusable, although far from fast. Maybe a better way of calculating $\vartheta$ could be found, or the GPU could be again usable for this task. We point to Nvidia's \cite{cusolver} as a possible starting point.
